### 2023-12-11 README.Rmd
Add Readme.md document.

---
output: github_document
bibliography: References.bib
---

<!-- *********** -->
<!-- Note: README.md is generated from README.Rmd.   -->
<!-- Be sure to edit README.Rmd and generate the README.md file by Cmd/Ctl-shift-K -->
<!-- *********** -->

<!-- Define some macros for later use -->
[link-cran]:             https://cran.r-project.org
[link-ecctools]:         https://github.com/earamendia/ECCTools
[link-eroitools]:        https://github.com/earamendia/EROITools
[link-iea]:              https://www.iea.org
[link-ieatools]:         https://github.com/MatthewHeun/IEATools
[link-ieaeweb]:          https://www.iea.org/data-and-statistics/data-product/world-energy-balances
[link-ilo]:              https://www.ilo.org/
[link-fao]:              https://www.fao.org/
[link-matsbyname]:       https://github.com/MatthewHeun/matsbyname
[link-matsindf]:         https://github.com/MatthewHeun/matsindf
[link-mwtools]:          https://github.com/EnergyEconomyDecoupling/MWTools
[link-pfusetup]:         https://github.com/EnergyEconomyDecoupling/PFUSetup
[link-pfudatabase]:      https://github.com/EnergyEconomyDecoupling/PFUDatabase
[link-pfuaggdatabase]:   https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
[link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
[link-pins]:             https://pins.rstudio.com
[link-rclabels]:         https://github.com/MatthewHeun/RCLabels
[link-recca]:            https://github.com/MatthewHeun/Recca
[link-targets]:          https://docs.ropensci.org/targets
[link-tidyverse]:        https://www.tidyverse.org

## JOSS paper

This repository contains a paper submitted to 
the [Journal of Open Source Software (JOSS)](https://joss.theoj.org)
covering the suite of `R` packages associated with the 
CL-PFU database^["CL-PFU" 
is an initialism for "Country-Level
Primary, Final, and Useful."
The CL-PFU database contains many countries as well as 
continental and world aggregations.
Technically speaking, 
we create [matsindf][link-matsindf] data frames
stored as [pins][link-pins] on a pinboard, 
not SQL or similar databases.].

## CL-PFU database

We created a suite of `R` packages 
to support the creation and use of the CL-PFU database.
The CL-PFU database is a new resouce 
for the Societal Exergy Analysis (SEA) research community and
assists SEA practitioners 
to analyze energy conversion chains (ECCs).
The new packages enable, for the first time, scalable SEA.

## `R` packages

The `R` packages covered by this paper include
three packages that support the PSUT framework of
[Heun et al. [-@Heun:2018]](https://doi.org/10.1016/j.apenergy.2018.05.109).
These packages are 
generally useful and available on [CRAN][link-cran].

| Package | Purpose |
|:--------|:--------|
| [RCLabels][link-rclabels]     | Manipulates matrix row and column names in [matsindf][link-matsindf] data frames |
| [matsbyname][link-matsbyname] | Performs matrix mathematics that respects row and column names |
| [matsindf][link-matsindf]     | Stores matrices in cells of data frames, thereby enabling analyses with [tidyverse][link-tidyverse] syntax |

Several additional packages support the creation of the CL-PFU database. 
The packages in the following table are available on GitHub.

| Package | Purpose |
|:--------|:--------|
| [IEATools][link-ieatools]             | Converts [IEA EWEB][link-ieaeweb] data to RUVY matrices in [matsindf][link-matsindf] data frames |
| [MWTools][link-mwtools]               | Converts [ILO][link-ilo] and [FAO][link-fao] data to RUVY matrices of human and animal muscle work in [matsindf][link-matsindf] data frames |
| [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
| [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
| [PFUPipelineTools][link-pfupipelinetools] | Provides basic functionality for all PFU pipelines |
| [PFUDatabase][link-pfudatabase]       | Provides a [targets][link-targets] pipeline to create a data frame of RUVY matrices |
| [PFUAggDatabase][link-pfuaggdatabase] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |

## References