### 2023-12-11 README.Rmd
Update README with statement of need.

---
output: github_document
bibliography: paper.bib
---

<!-- *********** -->
<!-- Note: README.md is generated from README.Rmd.   -->
<!-- Be sure to edit README.Rmd and generate the README.md file by Cmd/Ctl-shift-K -->
<!-- *********** -->

<!-- Define some macros for later use -->
[link-cran]:             https://cran.r-project.org
[link-ecctools]:         https://github.com/earamendia/ECCTools
[link-eroitools]:        https://github.com/earamendia/EROITools
[link-iea]:              https://www.iea.org
[link-ieatools]:         https://github.com/MatthewHeun/IEATools
[link-ieaeweb]:          https://www.iea.org/data-and-statistics/data-product/world-energy-balances
[link-ilo]:              https://www.ilo.org/
[link-fao]:              https://www.fao.org/
[link-matsbyname]:       https://github.com/MatthewHeun/matsbyname
[link-matsindf]:         https://github.com/MatthewHeun/matsindf
[link-mwtools]:          https://github.com/EnergyEconomyDecoupling/MWTools
[link-pfusetup]:         https://github.com/EnergyEconomyDecoupling/PFUSetup
[link-pfudatabase]:      https://github.com/EnergyEconomyDecoupling/PFUDatabase
[link-pfuaggdatabase]:   https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
[link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
[link-pins]:             https://pins.rstudio.com
[link-rclabels]:         https://github.com/MatthewHeun/RCLabels
[link-recca]:            https://github.com/MatthewHeun/Recca
[link-targets]:          https://docs.ropensci.org/targets
[link-tidyverse]:        https://www.tidyverse.org

## JOSS paper

This repository contains a paper submitted to 
the [Journal of Open Source Software (JOSS)](https://joss.theoj.org)
covering the suite of `R` packages associated with the 
CL-PFU^["CL-PFU" 
is an initialism for "Country-Level
Primary, Final, and Useful."
The CL-PFU database contains many countries as well as 
continental and world aggregations.
Technically speaking, 
we create [matsindf][link-matsindf] data frames
stored as [pins][link-pins] on a pinboard, 
not SQL or similar databases.] database.

## Statement of need

Historically, Societal Exergy Analysis (SEA)
practitioners have analyzed the ECCs
of individual countries using linked spreadsheets,
often starting with primary- and final-stage data 
from the [International Energy Agency][link-iea]'s (IEA's)
[extended world energy balances][link-ieaeweb] (EWEBs)
[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 
Data were stored in varying and inconsistent formats.
An early SEA database [@De-Stercke:2014]
estimated energy efficiencies 
of end-use machines using the economic status of countries,
thereby precluding use in energy-economy studies.

The authors of the current paper and others in the field
wanted to expand SEA to cover all countries, but
the spreadsheet approach to SEA was
deemed not scalable.
A new approach to SEA was needed, 
one that scaled across all countries 
for many years without relying on economic data
to estimate machine efficiencies.

## CL-PFU database

We created a suite of `R` packages 
to support the creation and use of the CL-PFU database.
The CL-PFU database is a new resouce 
for the SEA research community and
assists SEA practitioners 
to analyze energy conversion chains (ECCs).
The new packages enable, for the first time, scalable SEA.

## `R` packages

The `R` packages covered by this paper include
three packages that support the PSUT framework of
[Heun et al. [-@Heun:2018]](https://doi.org/10.1016/j.apenergy.2018.05.109).
These packages are 
generally useful and available on [CRAN][link-cran].

| Package | Purpose |
|:--------|:--------|
| [RCLabels][link-rclabels]     | Manipulates matrix row and column names in [matsindf][link-matsindf] data frames |
| [matsbyname][link-matsbyname] | Performs matrix mathematics that respects row and column names |
| [matsindf][link-matsindf]     | Stores matrices in cells of data frames, thereby enabling analyses with [tidyverse][link-tidyverse] syntax |

Several additional packages support the creation of the CL-PFU database. 
The packages in the following table are available on GitHub.

| Package | Purpose |
|:--------|:--------|
| [IEATools][link-ieatools]             | Converts [IEA EWEB][link-ieaeweb] data to RUVY matrices in [matsindf][link-matsindf] data frames |
| [MWTools][link-mwtools]               | Converts [ILO][link-ilo] and [FAO][link-fao] data to RUVY matrices of human and animal muscle work in [matsindf][link-matsindf] data frames |
| [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
| [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
| [PFUPipelineTools][link-pfupipelinetools] | Provides basic functionality for all PFU pipelines |
| [PFUDatabase][link-pfudatabase]       | Provides a [targets][link-targets] pipeline to create a data frame of RUVY matrices |
| [PFUAggDatabase][link-pfuaggdatabase] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |

## References